\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-06-07 nampi.tex
writing

The FlashFill system~\cite{gulwani2011automating} developed by
Microsoft researchers and now embedded in Excel solves problems such
as these and is probably the best known practical program-induction
algorithm, but researchers in programming languages and AI have built
successful program induction algorithms for many applications, 
such as handwriting recognition and generation~\cite{lake2015human},
procedural graphics~\cite{ellis2017learning}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}, to name just a few.
These systems work in different ways, but most hinge upon  a
carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
especially true for systems such as FlashFill that aim to induce a
wide range of programs very quickly, in a few seconds or less.  DSLs
constrain the search over programs with strong prior knowledge in the
form of a restricted set of programming primitives tuned to the needs
of the domain: for text editing, these are operations like appending
strings and splitting on characters.

In this work, we consider the problem of building agents that learn to solve
program induction tasks, and also the problem of acquiring the prior knowledge
necessary to quickly solve these tasks in a new domain.  Representative problems
in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
algorithm that grows or boostraps a DSL while jointly training a neural network
to help write programs in the increasingly rich DSL.
%
% (JBT: MAYBE INSERT FIGURE FROM ICML PAPER?)
%
\begin{table*} %[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=4pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

    \end{tabular}
    ~\\
    \midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \hspace{0cm}\begin{tabular}{l}
      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
      %NOTE: this is the actual invention, but I removed a redundant lambda
      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
      \phantom{$f_1(\ell,$}
      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
      \hspace{\helpSize}($f_3$: \emph{Whether $\ell$ contains }\code{k})\\
    \end{tabular}&

  \hspace{0.5cm}\begin{tabular}{l}
    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
    (x)}}\\
    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
      }}\\
      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
        \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
      \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
  \end{tabular}&

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
\end{table*}
Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, to learn to synthesize large bodies of code, or to learn
complex programs starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that humans can solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  Even with a good DSL, program search may be
intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
so we train a
neural network to guide the search procedure.

@@

We take inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on characters such as spaces or commas that are
frequently used to delimit substrings.  Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution.%% ; if every output
%% contains a constant substring (e.g., ``Dr.''), inserting or appending
%% that constant string is likely to be a subroutine.

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.

@@

\section{Programs that manipulate sequences}\label{sequences}
We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for
the recognition model, and initially provide the system with generic list processing primitives:
\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{List Processing}\label{listSection}
Synthesizing programs that manipulate data structures is a widely studied
problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of learning functions that
manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
\begin{figure}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}
\end{figure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples,
%% Our data set is interesting in three major ways: many of the tasks
%% require complex solutions; the tasks were not generated from some
%% latent DSL, and the agent must learn to solve these complicated
%% problems from only 236 tasks.
%% Our data set assumes arithmetic operations as well as sequence operations,
%% so we additionally provide our system with the following arithmetic
%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
%% \code{is-prime}.
evaluating  on a 50/50 test/train split.  The system
composed 38 new subroutines, and rediscovered the higher-order
function \code{filter}, yielding a more expressive DSL more closely
matching the domain (Tbl.~\ref{initialExampleDSL}, left).

\subsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{lau2001programming},
and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
We show \system can instead start out with generic sequence manipulation
primitives and recover many of the higher-level building blocks that have
made these other text editing systems successful.

%% Because our enumerative search procedure cannot generate string %
%% constants, we instead enumerate programs with string-valued
%% parameters.  For example, to learn a program that prepends ``Dr.'', we
%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
%% define $\probability[x|p]$ by approximately marginalizing out the
%% string parameters via a simple dynamic program.
%% In Sec.~\ref{regressionSection}, we will use a similar trick to
%% synthesize programs containing real numbers, but using gradient
%% descent instead of dynamic programming.

@@

\section{Quantitative Results}\label{quantitative}

@@

\begin{figure*}\centering
  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
  \caption{Learning curves for \system both with (\orange{in orange}) and without
  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time, averaged only over those tasks that are solved.}\label{learningCurves}
\end{figure*}

@@

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration of top-down
symbolic representations and bottom-up neural networks --- both of them learned
--- helps make program induction systems more generally useful for AI. Many
directions remain open.
%\paragraph{Future.}
Two immediate goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
Another direction is to explore DSL meta-learning: can we find a
\emph{single} universal primitive set that could effectively bootstrap DSLs for
new domains, including the three domains considered,  but also many others?

>>>>

The FlashFill system~\cite{gulwani2011automating} embedded in
Microsoft Excel solves problems such as these and is probably the best
known practical program-induction algorithm, but researchers in
programming languages and AI have built successful program induction
algorithms for many applications, such as handwriting recognition and
generation~\cite{lake2015human}, procedural
graphics~\cite{ellis2017learning}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}.  These systems
work in different ways, but most hinge upon a carefully engineered
\textbf{Domain Specific Language (DSL)}.  This is especially true for
systems such as FlashFill that aim to induce a wide range of programs
very quickly, in a few seconds or less.  DSLs constrain the search
over programs with strong prior knowledge in the form of a restricted
set of programming primitives tuned to the needs of the domain: for
text editing, these are operations like appending and
splitting on characters.

In this work, we consider the problem of building agents that learn to solve
program induction tasks, and also the problem of acquiring the prior knowledge
necessary to quickly solve these tasks in a new domain.  Representative problems
in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
algorithm that grows or boostraps a DSL while jointly training a neural network
to help write programs in the increasingly rich DSL.
%
% (JBT: MAYBE INSERT FIGURE FROM ICML PAPER?)
%
\begin{table*} %[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=4pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

    \end{tabular}
    ~\\
    \midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \hspace{0cm}\begin{tabular}{l}
      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
      %NOTE: this is the actual invention, but I removed a redundant lambda
      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
      %% \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
      %% \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
      \phantom{$f_1(\ell,$}
      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
      \hspace{\helpSize}($f_3$: \emph{Whether $\ell$ contains }\code{k})\\
    \end{tabular}&

  \hspace{0.5cm}\begin{tabular}{l}
    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
    (x)}}\\
    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
      }}\\
      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
      %%   \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
      %% \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
  \end{tabular}&

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
\end{table*}
Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that people solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  %% Even with a good DSL, program search may be
%% intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
%% so we train a
%% neural network to guide the search procedure.

@@

We take inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on frequent delimiters such as spaces or commas. Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution.%% ; if every output
%% contains a constant substring (e.g., ``Dr.''), inserting or appending
%% that constant string is likely to be a subroutine.

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.

@@

\section{Programs that manipulate sequences}\label{sequences}
We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}),
using a bidirectional GRU~\cite{cho2014learning} for
the recognition model, and initially providing the system with generic sequence manipulation primitives:
\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{List Processing}\label{listSection}
Synthesizing programs that manipulate data structures is a widely studied
problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of learning functions that
manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
\begin{figure}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
%    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
%    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}
\end{figure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples.
%% Our data set is interesting in three major ways: many of the tasks
%% require complex solutions; the tasks were not generated from some
%% latent DSL, and the agent must learn to solve these complicated
%% problems from only 236 tasks.
%% Our data set assumes arithmetic operations as well as sequence operations,
%% so we additionally provide our system with the following arithmetic
%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
%% \code{is-prime}.
%% evaluating  on a 50/50 test/train split.
In solving these tasks, the system
composed 38 new subroutines, and rediscovered the higher-order
function \code{filter} (Tbl.~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).

\subsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{lau2001programming},
and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
Here we will instead start out with generic sequence manipulation
primitives and recover many of the higher-level building blocks that have
made these other text editing systems successful.
%% Because our enumerative search procedure cannot generate string %
%% constants, we instead enumerate programs with string-valued
%% parameters.  For example, to learn a program that prepends ``Dr.'', we
%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
%% define $\probability[x|p]$ by approximately marginalizing out the
%% string parameters via a simple dynamic program.
%% In Sec.~\ref{regressionSection}, we will use a similar trick to
%% synthesize programs containing real numbers, but using gradient
%% descent instead of dynamic programming.

@@

\section{Quantitative Results}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them~\ref{learningCurves}.
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models that
more closely match the domains.

@@

\begin{figure*}\centering
  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
  \caption{Learning curves for \system both with (\orange{in orange}) and without
  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved w/ 10m timeout. Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}
\end{figure*}

@@

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration of top-down
symbolic representations and bottom-up neural networks --- both of them learned
--- can help make program induction systems more generally useful for AI. Many
directions remain open.
%\paragraph{Future.}
Two immediate goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
Another direction is to explore DSL meta-learning: can we find a
\emph{single} universal primitive set that could effectively bootstrap DSLs for
new domains, including the three domains considered,  but also many others?
